---
title: "WEEK14 ASSOCIATION RULES"
author: "Joseph Kiburu"
date: "2/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# We first we install the required arules library 
library(arules)
```

```{R}
# Loading our dataset

df<-read.transactions('http://bit.ly/SupermarketDatasetII', sep=",")
head(df)

```

```{r}
# Verifying the object's class

class(df)
```

```{r}
# Previewing our first 5 transactions
inspect(df[1:5])
```

```{r}
# If we wanted to preview the items that make up our dataset,
# alternatively we can do the following
# ---
# 
dataset<-as.data.frame(itemLabels(df))
colnames(dataset) <- "Item"
head(dataset, 10) 

```


```{r}
# Summary of the transaction dataset

summary(df)
```

Mineral water is the most frequent item being bought.

```{r}
# Checking the frequency of items ranging from 5 to 8.
itemFrequency(df[, 5:8],type = "absolute")
round(itemFrequency(df[, 5:8],type = "relative")*100,2)
```

From range 5 to 8, black tea has the highest number of transactions while babies food has the least.

```{r}
# Visualizing top 10 most common items in our dataset and the items whose relative importance is at least 10%
par(mfrow = c(1, 2))

# plotting the frequency of items
itemFrequencyPlot(df, topN = 10,col="darkgreen")
itemFrequencyPlot(df, support = 0.1,col="darkred")

```

From our 10% support range, Mineral water has the highest support while milk has the least. Because of its popularity, there is a high number of mineral water being bought.

```{r}
# Building a model based on association rules using the apriori function using a 
# support of 0.002 and confidence as 0.7.


rules <- apriori (df, parameter = list(supp = 0.002, conf = 0.7))
rules


```


Using a higher level of support can make the model lose interesting rules. In our case, we have 11 rules from a support of 0.002. Using a low confidence level increases the number of rules.

```{r}
# Checking the information about our model.
summary(rules)

```


Most rules have 3 and 4 itemS with a support of 0.002 and a 70% confidence.

```{r}
# Observing rules built in our model
# Checking our first six model rules.

inspect(rules)
```

For the first rule, if someone buys nonfat milk and spaghetti, they are 72% likely to buy mineral water. Of all the rules, rule two is the best rule with a confidence of 95%.

```{r}
# Ordering these rules by the level of confidence

rules<-sort(rules, by="confidence", decreasing=TRUE)
inspect(rules)
```

Rules arranged from the best to the worst using the level of confidence as the criteria.